---
title: "Akbari_Peak_Bloom_Pred"
author: "Matin Akbari"
date: "2/8/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
We need to predict doy for Kyoto (Japan), Liestal-Weideli (Switzerland), Washington D.C. (USA), Vancouver BC (Canada)

1) Lets begin with some research on cherry trees and what makes them bloom. 

We need location, Latitude, Longitude, Altitude, Date, doy, and Temperture for each of the locations (minus Vancouver)


```{r}
#packages used
library(tidyverse)
library(stringr)
library(rnoaa)
library(mice)
library(VIM)
library(forecast)
```

```{r}
#Getting the dataset for historic cherry tree bloom data
cherry <- read.csv("peak-bloom-prediction/data/washingtondc.csv") %>% 
  bind_rows(read.csv("peak-bloom-prediction/data/liestal.csv")) %>% 
  bind_rows(read.csv("peak-bloom-prediction/data/kyoto.csv"))
```
We are going to use data from 1950 and onward. 

First we need to do some feature engineering to get the dataset to look how we want.

```{r}
cherry2 = rename(cherry, "bloom_year" = "year")
cherry2$bloom_year = as.integer(str_sub(cherry$bloom_date, 1,4))
cherry2$bloom_month = as.integer(str_sub(cherry$bloom_date, 6,7))
cherry2$bloom_day = as.integer(str_sub(cherry$bloom_date, 9))
cherry2 = cherry2[,c(1:4,6,8,9,5,7)]
```

```{r}
#Getting historic temperture using function from demo
stations <- ghcnd_stations()

get_temperature <- function (stationid) {
  ghcnd_search(stationid = stationid, var = c("tmax"), 
               date_min = "1880-01-01", date_max = "2022-01-31")[[1]] %>%
  mutate(year = as.integer(format(date, "%Y")),
         month = as.integer(strftime(date, '%m')) %% 12, # make December "0"
         season = cut(month, breaks = c(0, 2, 5, 8, 11),
                      include.lowest = TRUE,
                      labels = c("Winter", "Spring", "Summer", "Fall")),
         year = if_else(month == 0, year + 1L, year)) %>%
  group_by(year, season) %>%
  summarize(tmax_avg = mean(tmax, na.rm = TRUE))
}

historic_temperatures <-
  tibble(location = "washingtondc", get_temperature("USC00186350")) %>%
  bind_rows(tibble(location = "liestal", get_temperature("GME00127786"))) %>%
  bind_rows(tibble(location = "kyoto", get_temperature("JA000047759"))) %>%
  bind_rows(tibble(location = "vancouver", get_temperature("CA001108395")))

historic_temperatures %>%
  ggplot() + 
  aes(year, tmax_avg) + 
  geom_line() +
  xlim(1950, 2031) +
  labs(x = "Year", y = "Average maximum temperature (1/10 °C)") +
  facet_grid(factor(season) ~ str_to_title(location))

temp_dat = historic_temperatures[historic_temperatures$season=="Spring",-3]
temp_dat = temp_dat[!(temp_dat$location=="vancouver"),]
```

```{r} 
#merging temp with rest of data
full_cherry = merge(cherry2, temp_dat, by.x = c("location","bloom_year"), by.y = c("location","year"), all.x = TRUE, all.y = TRUE)
full_cherry = full_cherry[,c(1:5,10,6:9)]
full_cherry = full_cherry[full_cherry$bloom_year>=1950,]
full_cherry$location = as.factor(full_cherry$location)
```

```{r}
#imputing missing data
matrixplot(full_cherry)
VIM::aggr(full_cherry)

fcs = mice(full_cherry[,c(1:6,8:10)], m=5, method="pmm")
train_comp = complete(fcs,2)
full_cherry$tmax_avg = train_comp$tmax_avg
VIM::aggr(full_cherry)
```

```{r}
#visualization of data
plot(full_cherry$bloom_doy~full_cherry$tmax_avg)

```

```{r}
#using ARIMA to forecast predictions for next decade
nrow(full_cherry)
full_cherry = full_cherry[order(full_cherry$location),-c(7)]

kyoto_model = full_cherry[full_cherry$location=="kyoto",]
fit_ky_arima = arima(kyoto_model[,9])
preds_ky = forecast(fit_ky_arima,h=10)

liestal_model = full_cherry[full_cherry$location=="liestal",]
fit_li_arima = arima(liestal_model[,9]) 
preds_li = forecast(fit_li_arima,h=10)

washingtondc_model = full_cherry[full_cherry$location=="washingtondc",]
fit_wa_arima = arima(washingtondc_model[,9]) 
preds_wa = forecast(fit_wa_arima, h=10)

preds_final = data.frame(year=c(2022,2023,2024,2025,2026,2027,2028,2029,2030,2031),kyoto=preds_ky$mean,liestal=preds_li$mean, washingtondc=preds_wa$mean)
preds_final$kyoto = round(preds_final$kyoto)
preds_final$liestal = round(preds_final$liestal)
preds_final$washingtondc = round(preds_final$washingtondc)
#writing a loop to take the combined average from the other sites to forecast vancouver
preds_final$vancouver = NA
for (i in 1:nrow(preds_final)) {
  preds_final$vancouver[i] = round((preds_final$kyoto[i]+preds_final$liestal[i]+preds_final$washingtondc[i])/3)
}
write.csv(preds_final, file = "C:\\Users\\admin\\Desktop\\Course Work\\Peak Bloom Predictions\\Akbari_Peak_Bloom_Pred\\Akbari_Predictions.txt", row.names = FALSE)
```
